---
title: "Sleep Health and Lifestyle Analysis - Final Project"
author: "Zixuan Li"
date: "`r Sys.Date()`"
output: html_document
params:
  production: TRUE
---

```{r, set-chunk-opts, echo = FALSE, eval = params$production}
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE
)
```


# Final Project Introduction
The Sleep Health and Lifestyle Dataset aim to provide detailed insights into the sleep patterns, lifestyle behaviors, and cardiovascular health of 400 individuals, covering 13 variables such as age, gender, sleep duration, quality of sleep, physical activity levels, stress levels, BMI category, blood pressure, heart rate, daily steps, and sleep disorders.

  1. **Person ID**: An identifier for each individual.
  2. **Gender**: The gender of the person (Male/Female).
  3. **Age**: The age of the person in years.
  4. **Occupation**: The occupation or profession of the person.
  5. **Sleep Duration (hours)**: The number of hours the person sleeps per day.
  6. **Quality of Sleep**: A subjective rating of the quality of sleep, ranging from 1 to 10.
  7. **Physical Activity Level** (minutes/day): The number of minutes the person engages in physical activity daily.
  8. **Stress Level**: A subjective rating of the stress level experienced by the person, ranging from 1 to 10.
  9. **BMI Category**: The BMI category of the person (e.g., Underweight, Normal, Overweight).
  10. **Blood Pressure** (systolic/diastolic): The blood pressure measurement of the person, indicated as systolic pressure over diastolic pressure.
  11. **Heart Rate** (bpm): The resting heart rate of the person in beats per minute.
  12. **Daily Steps**: The number of steps the person takes per day.
  13. **Sleep Disorder**: The presence or absence of a sleep disorder in the person (None, Insomnia, Sleep Apnea).
      * None: The individual does not exhibit any specific sleep disorder.
      * Insomnia: The individual experiences difficulty falling asleep or staying asleep, leading to inadequate or poor-quality sleep.
      * Sleep Apnea: The individual suffers from pauses in breathing during sleep, resulting in disrupted sleep patterns and potential health risks.
  
**Objective:** The objective of this analysis is for in-depth exploration of the relationships between sleep quality and various health and lifestyle factors, helping to identify key drivers of healthy sleep and overall well-being. Specifically, I aim to identify key factors that influence sleep quality.

Data source:
https://www.kaggle.com/datasets/uom190346a/sleep-health-and-lifestyle-dataset/discussion/426177


```{r, here-i-am, include = FALSE}
here::i_am(
  "final_report.Rmd"
)
```

```{r, load-data}
data <- readRDS(
  file = here::here("data/data_cleaned.rds")
)
```

# Descriptive Analysis

## Tabular analysis

##### Table 1: Descriptive summary table of key variables stratified by Sleep Quality

1. The table provides a descriptive summary of key variables stratified by Quality of Sleep (which has categories like "Poor-Moderate" and "Good").

2. Continuous variables are summarized as mean (SD), giving insights into the distribution of variables such as *Age*, *Sleep Duration*, and *Physical Activity Level* across different levels of sleep quality.

3. Categorical variables like *Gender*, *BMI Category*, and *Sleep Disorder*, the table reports counts and percentages as n(%).

4. An additional overall column is included to show the summary for the entire dataset.

5. P-values assess whether there are statistically significant differences between the different levels of sleep quality for each variable.

```{r, table1}
table1 <- readRDS(
  file = here::here("output/table1.rds")
)
table1
```

## Graphical analysis

##### Figure 1: Distribution of Sleep Duration by Quality of Sleep

This box plot depicts the distribution of Sleep Duration across different Quality of Sleep categories (Poor-Moderate and Good). The median sleep duration is represented by the line inside each box, with the boxes displaying the IQR.

```{r, plots}
#| fig.align = "center",
#| out.width = "600px"
knitr::include_graphics(
  here::here("output/plot1_box.png")
)
```

##### Figure 2: Proportion of Sleep Disorder by Quality of Sleep

This stacked bar plot illustrates the relationship between Sleep Disorder categories (None, Insomnia, Sleep Apnea) and Quality of Sleep (Poor-Moderate and Good). The height of each segment in the bars represents the proportion of individuals within each sleep disorder category for the respective quality of sleep level.

```{r}
#| fig.align = "center",
#| out.width = "600px"
knitr::include_graphics(
  here::here("output/plot2_bar.png")
)
```

